Continuous time

### 2023-09-25 continuous-time.qmd
cleaning

library(scales)
library(cmdstanr)
library(posterior)
library(brms)
library(ggstance)
library(ggdist)
library(patchwork)
library(tidyverse)
library(splines)
library(emmeans)
library(lme4)
library(broom.mixed)
library(knitr)
source("functions.R")
source("R/fit_cached.R")
```

@@

## Data
Load the data and remove observations that occurred more than 3 hours into a session.
```{r}
#| label: load-data
dat <- read_csv(
  "data/study_prompt_answered.csv",
  col_select = c(
    pid,
    time = Time_utc,
    duration = CurrentSessionLength,
    prompt = LastStudyPromptType,
    mood = response
  )
) |>
  mutate(
    pid = factor(pid),
    mood = mood / 1000
  )

# Create sessions using heuristics below
dat <- dat |>
  arrange(pid, time) |>
  mutate(
    i = row_number(),
    # Observation is in a new session if this...
    new_session =
    # is the first total observation,
      i == 1 |
        # session duration is zero,
        (duration == 0) |
        # session is shorter than previous session (ticker has reset),
        (duration < lag(duration, default = 0)) |
        # or if more than 30 minutes since previous observation.
        (time > lag(time) + minutes(30)),
    # Session number is the cumulative sum of new sessions...
    session = cumsum(new_session),
    # ... per person
    .by = pid
  ) |>
  select(-c(i, new_session))

dat <- dat |>
  mutate(
    duration = duration - duration[1],
    hours = duration / 60,
    .by = c(pid, session)
  ) |>
  mutate(
    ps = interaction(pid, session, sep = "_")
  ) |>
  filter(prompt == "Wellbeing")

@@

```{r}
#| label: tbl-session-max-prop
#| tbl-cap: Proportion of participants that had at least on session last longer than gt_* hours
dat |> 
  summarize(
    max = max(hours), .by = c(pid)
  ) |> 
  summarise(
    gt_0.33 = mean(max > 0.33),
    gt_0.71 = mean(max > 0.716),
    gt_1 = mean(max > 1),
    gt_2 = mean(max > 2),
  ) |> 
  kable(digits = 2)
```

@@

```{r}
#| label: define-spline
ns4 <- function(x) {
  ns(
    x,
    knots = c(0.05, 0.33, 0.7),
    Boundary.knots = c(0, 3)
  )
}
```

@@

```{r}
#| label: lmm-ns4
fit <- fit_cached(
  "models/lmm-ns4.Rds",
  lmer(
    mood100 ~ ns4(hours) + (1 | ps) + (1 + ns4(hours) | pid),
    data = dat
  )
)
```

@@

```{r}
#| label: fig-session-change
#| echo: false
#| fig-width: 8
#| fig-height: 5
p_comb <- (p_curve) /
  (
    p_int_sd +
      p_int_sd_change +
      plot_layout(guides = "collect") &
      theme(legend.position = "bottom")
  ) +
  plot_annotation(tag_levels = "A")

p_comb
ggsave(
  "figures/fig-session-change.png",
  p_comb,
  width = 8,
  height = 5.5
)
```

@@

```{r}
#| label: lmer-pre-interaction
dat_pre <- dat |>
  filter(
    # Session has a wellbeing measure at time = 0
    duration[1] == 0,
    .by = c(pid, session)
  ) |>
  mutate(
    pre = mood100[1],
    .by = c(pid, session)
  ) |>
  filter(
    hours != 0
  )

@@

```{r}
#| label: pre-sum
pre_sum <- dat_pre |>
  summarise(
    pre_mean = median(pre),
    .by = pid
  ) |>
  summarise(
    mean = mean(pre_mean),
    sd = sd(pre_mean)
  )
kable(pre_sum, digits = 2)
```

```{r}
#| label: pre-sd-values
pre_values <- pre_sum$mean + pre_sds * pre_sum$sd
data.frame(
  pre = pre_values,
  label = pre_labels
  ) |> 
  kable(digits = 2)
```

@@

### Manuscript figures
Create figures for the manuscript.
```{r}
#| label: fig-ms-interaction
#| fig-height: 4
#| fig-width: 8

@@

```{r}
#| label: fig-ms-no-interaction
#| fig-height: 4
#| fig-width: 8
p_tmp <- (
  p_curve + labs(x = NULL, y = "Mood")
  ) / (
  p_change + labs(x = "Duration", y = "Change")
  ) + plot_annotation(tag_levels = "A") 
p_tmp
```

@@

The main model used a censored normal distribution. A similar model would take a long time to fit here, so we'll refit the continuous time model using a censored normal distribution on a random subset of 1000 participants.

```{r}
#| label: brms-cens
set.seed(345345)
tmp <- filter(
  dat,
  pid %in% sample(unique(pid), 1000)
)

@@

Lastly, we'll also fit the interaction model using a censored normal distribution, again with 1000 participants.

```{r}
#| label: cens-pre-interaction
set.seed(5454545)
tmp_pre <- filter(dat_pre, pid %in% sample(unique(pid), 1000))

@@

Next, we'll lok at `hours`. In @fig-lmm-pre-interaction-sens-hours we see that the results aren't fundamentally different using splines with less degrees of freedom, or even a linear term.

>>>>

library(scales)
library(cmdstanr)
library(posterior)
library(brms)
library(patchwork)
library(tidyverse)
library(splines)
library(emmeans)
library(lme4)
library(broom.mixed)
library(knitr)
source("functions.R")
source("R/fit_cached.R")
```

@@

## Data
Load the data and remove observations that occurred more than 3 hours into a session.
```{r}
#| label: load-data
dat <- read_rds("data/data-raw.rds")
if (!is.na(N_SUBSET)) {
  pid_subset <- sample(unique(dat$pid), N_SUBSET)
  dat <- filter(dat, pid %in% pid_subset)
}
dat <- dat |>
  mutate(
    duration = duration - duration[1],
    hours = duration / 60,
    .by = c(pid, session)
  ) |>
  mutate(
    ps = interaction(pid, session, sep = "_")
  ) |>
  filter(prompt == "Wellbeing")

@@

```{r}
#| label: tbl-session-max-prop
#| tbl-cap: Proportion of participants that had at least on session last longer than gt_* hours
dat |> 
  summarise(
    max = max(hours), .by = c(pid)
  ) |> 
  summarise(
    gt_0.33 = mean(max > 0.33),
    gt_0.71 = mean(max > 0.716),
    gt_1 = mean(max > 1),
    gt_2 = mean(max > 2),
  ) |> 
  kable(digits = 2)
```

@@

```{r}
#| label: define-spline
#| code-fold: show
ns4 <- function(x) {
  ns(
    x,
    knots = c(0.05, 0.33, 0.7),
    Boundary.knots = c(0, 3)
  )
}
```

@@

```{r}
#| label: lmm-ns4
#| code-fold: show
fit <- fit_cached(
  "models/lmm-ns4.Rds",
  lmer(
    mood100 ~ ns4(hours) + (1 | ps) + (1 + ns4(hours) | pid),
    data = dat
  )
)
```

@@

@@

```{r}
#| label: lmer-pre-interaction
#| code-fold: show
dat_pre <- dat |>
  filter(
    # Session has a wellbeing measure at time = 0
    duration[1] == 0,
    .by = c(pid, session)
  ) |>
  mutate(
    pre = mood100[1],
    .by = c(pid, session)
  ) |>
  filter(
    hours != 0
  )

@@

```{r}
#| label: tbl-pre-sum
#| tbl-cap: Average mood levels the start of a session
pre_sum <- dat_pre |>
  summarise(
    pre_mean = median(pre),
    .by = pid
  ) |>
  summarise(
    mean = mean(pre_mean),
    sd = sd(pre_mean)
  )
kable(pre_sum, digits = 2)
```

```{r}
#| label: tbl-pre-sd-values
#| tbl-cap: Distribution of average mood levels at the start of a session
pre_values <- pre_sum$mean + pre_sds * pre_sum$sd
data.frame(
  pre = pre_values,
  label = pre_labels
  ) |> 
  kable(digits = 2)
```

@@

### Manuscript figures
Create figures for the manuscript.
```{r}
#| label: fig-ms-interaction
#| fig-height: 4
#| fig-width: 8
#| fig-cap: Figure included in manuscript

@@

```{r}
#| label: fig-ms-no-interaction
#| fig-height: 4
#| fig-width: 8
#| fig-cap: Figure shown in manuscript
p_tmp <- (
  p_curve + labs(x = NULL, y = "Mood")
  ) / (
  p_change + labs(x = "Duration", y = "Change")
  ) + plot_annotation(tag_levels = "A") 
p_tmp
```

@@

```{r}
#| label: n-pids-sample
#| echo: false
n_pids_sample <- case_when(
  is.na(N_SUBSET) ~ 1000,
  !is.na(N_SUBSET) ~ min(N_SUBSET, 1000)
)
```

The main discrete time model used a censored normal distribution. A similar model would take a long time to fit here, so we'll refit the continuous time model using a censored normal distribution on a random subset of `r n_pids_sample` participants.

```{r}
#| label: brms-cens
set.seed(345345)
tmp <- filter(
  dat,
  pid %in% sample(unique(pid), n_pids_sample)
)

@@

Lastly, we'll also fit the interaction model using a censored normal distribution, again with `r n_pids_sample` participants.

```{r}
#| label: cens-pre-interaction
set.seed(5454545)
tmp_pre <- filter(dat_pre, pid %in% sample(unique(pid), n_pids_sample))

@@

Next, we'll look at `hours`. In @fig-lmm-pre-interaction-sens-hours we see that the results aren't fundamentally different using splines with less degrees of freedom, or even a linear term.
